Performance-portable Parallel Programming in Computer Science Education with Kokkos

### 2020-01-23 src/1.Introduction.tex
The David Poliakoff Center For Kids Who Want To Write Good

Kokkos\footnote{from Greek for grain or seed}~\cite{KOKKOS_PAPER_HERE}, is a performance-portable, parallel programming model for C++ developed at Sandia National Laboratories (SNL) for developing million line scientific software on the world\'s biggest supercomputers. It consists of a programming model specification and a library. The specification defines execution and memory model and an application programming interface (library API). The runtime library exposes functions representing parallel patterns and types for data management and maps them to the underlying abstract execution model. This execution model creates a generic view of compute hardware and memories. The Kokkos library maps execution primitives to vendor-specific programming models, libraries, and memory layouts. Kokkos has been successfully used in many large-scale projects at SNL, LLNL and other research centers~\cite{CITEKOKKOSUSECASES}. Its key properties are

@@

First, Kokkos has semantics that are generic. This means that students are learning parallelism detached from any architectural concerns. Kokkos won\'t require them to know the specifics of the \"symmetric multiprocessors\" of a GPU or the cache hiearchy on a CPU\footnote{Though nothing precludes an educator from teaching these independently of the model}. Finally, Kokkos is vendor agnostic. If you have a computer lab with processors from one vendor and your students have processors from another, this presents no problem, Kokkos likely works on both. Kokkos is declarative, students aren\'t saying \"map these work items to this set of threads\" but \"do a reduction over this set of data.\" This is critical, much of software engineering is moving to declarative models and languages, learning to operate in such a mindset is useful outside of parallel programming.

>>>>

Kokkos\footnote{from Greek for grain or seed}~\cite{KOKKOS_PAPER_HERE}, is a performance-portable, parallel programming model for C++ developed at Sandia National Laboratories (SNL) for developing million line scientific software on the world's biggest supercomputers. It consists of a programming model specification and a library. The specification defines execution and memory model and an application programming interface (library API). The runtime library exposes functions representing parallel patterns and types for data management and maps them to the underlying abstract execution model. This execution model creates a generic view of compute hardware and memories. The Kokkos library maps execution primitives to vendor-specific programming models, libraries, and memory layouts. Kokkos has been successfully used in many large-scale projects at SNL, LLNL and other research centers~\cite{CITEKOKKOSUSECASES}. Its key properties are

@@

Kokkos has semantics that are generic. This means that students are learning parallelism detached from any architectural concerns. Kokkos won't require them to know the specifics of the "symmetric multiprocessors" of a GPU or the cache hiearchy on a CPU\footnote{Though nothing precludes an educator from teaching these independently of the model}. Kokkos is vendor agnostic. If you have a computer lab with processors from one vendor and your students have processors from another, this presents no problem, Kokkos likely works on both. Kokkos is declarative, students aren't saying "map these work items to this set of threads" but "do a reduction over this set of data." This is critical, much of software engineering is moving to declarative models and languages, learning to operate in such a mindset is useful outside of parallel programming. All of these are desirable in a programming model for education.
